Global Warming Distributed Data Analysis

### 2016-03-28 report/article.tex
Conclusions e mappa che mostra alzamento temperature

\subsection{Temperature data quality}
\label{sec:filter}
Before using the actual temperature data, we filtered it through a custom quality check. Each measurement come with a quality flag set at a different value depending on how reliable the measurement was\cite{noaaDataFormat}.
It is important to use this flag to remove wrong data in the dataset, as reported in the a document provided by NOAA that describes the problem inside the dataset \cite{noaaProblems}. ``Various, mostly random errors, are present in the overall dataset, and these are being documented and scheduled for correction. Quality control of the data
(which has already taken place and reflected in the online data) has corrected
well over 99 percent of the errors present in the original data prior to the quality control.''
In particular we accepted all the measurement with the following quality code value:
\begin{itemize}
    \item 0: Passed gross limits check;
    \item 1: Passed all quality control checks;
    \item 4: Passed gross limits check, data originate from an NCEI data source;
    \item 5: Passed all quality control checks, data originate from an NCEI data source;
    \item A: Data value flagged as suspect, but accepted as a good value;
    \item I: Data value not originally in data, but inserted by validator;
    \item M: Manual changes made to value based on information provided by NWS or FAA;
    \item P: Data value not originally flagged as suspect, but replaced by validator;
    \item R: Data value replaced with value computed by NCEI software;
    \item U: Data value replaced with edited value.
\end{itemize}
We opted for a loose policy on the data quality, accepting also values that didn't pass all the quality checks or that were manually or software replaced. We refuted all the codes that stand for erroneous, missing or suspect data that was not manually or automatically checked and replaced by the institution. Furthermore, we applied a bound check for all the values greater than the maximum registered temperature (+61.8 \celsius) and the minimum one (-93.2 \celsius). \\
The data collected can be of two different types depending on the origin: from a fixed weather station or from a mobile one. In the second case, it's possible that the data registered from a mobile station (e.g. on a ship) is reported at a certain position and the following measurement is reported at a different position. Since we didn't distinguish data by time of the day, it's important that the number of measurements from a certain position is approximately the same for different hours. Since this is not guaranteed by moving stations, we decided to filter the measurement obtained from this source. \\

@@

In the graphic interface there is the possibility to select the two set of five-year results to compare. The heatmap shows a colored square in a latitude and longitude quadrant if it's present a data for that quadrant in both the two sets of results. Till the period 1921-1925 there are few data in the results, with consequently few colored squares in the map. In the interface there is also the possibility to change the temperature metric that is used to compared the two set of results.\\
One problem of the interface is that the colored squares used in the map should change their height depending on the latitude value where they are positioned because the map projection isn't equiareal. The consequence in the heatmap visualization is that there is some empty space between the squares of a certain latitude value and the squares with the value of the latitude increased or decreased by one. This aspect makes the heatmap looks like as it is missing some data for certain latitude when instead the data are present, just the squares should be rectangles with a bigger height value.

@@

\section{Conclusions}
\label{sec:con}
Before answering the question we asked earlier we will take a brief general overview of the characteristics of this project. One of the main problem we encountered is the limited (if not absence) of documentation describing the data. Especially for the older datasets we had to directly analyze the data to know the origin (region on earth) or the time coverage (day/night reports, coverage of all months). Moreover, we had to take into account during the analysis and the result discussion the anomalies we discovered in the station number charts (\ref{App:AppendixA}). Without a uniform and well formed source it's more difficult to extract reliable statistics.\\

>>>>

\subsection{Temperature data quality}
\label{sec:filter}
Before using the actual temperature data, we filtered it through a custom quality check. Each measurement come with a quality flag set at a different value depending on how reliable the measurement was\cite{noaaDataFormat}.
It is important to use this flag to remove wrong data in the dataset, as reported in the a document provided by NOAA that describes the problem inside the dataset \cite{noaaProblems}: ``Various, mostly random errors, are present in the overall dataset, and these are being documented and scheduled for correction. Quality control of the data
(which has already taken place and reflected in the online data) has corrected
well over 99 percent of the errors present in the original data prior to the quality control.''
In particular we accepted all the measurement with the following quality code value:
\begin{itemize}
    \item 0: Passed gross limits check;
    \item 1: Passed all quality control checks;
    \item 4: Passed gross limits check, data originate from an NCEI data source;
    \item 5: Passed all quality control checks, data originate from an NCEI data source;
    \item A: Data value flagged as suspect, but accepted as a good value;
    \item I: Data value not originally in data, but inserted by validator;
    \item M: Manual changes made to value based on information provided by NWS or FAA;
    \item P: Data value not originally flagged as suspect, but replaced by validator;
    \item R: Data value replaced with value computed by NCEI software;
    \item U: Data value replaced with edited value.
\end{itemize}
We opted for a loose policy on the data quality, accepting also values that didn't pass all the quality checks or that were manually or software replaced. We refuted all the codes that stand for erroneous, missing or suspect data that was not manually or automatically checked and replaced by the institution. Furthermore, we applied a bound check for all the values greater than the maximum registered temperature (+61.8 \celsius) and the minimum one (-93.2 \celsius). \\
The data collected can be of two different types depending on the origin: from a fixed weather station or from a mobile one. In the second case, it's possible that the data registered from a mobile station (e.g. on a ship) is reported at a certain position and the following measurement is reported at a different position. Since we didn't distinguish data by time of the day, it's important that the number of measurements from a certain position is approximately the same for different hours. Since this is not guaranteed by moving stations, we decided to filter the measurement obtained from this source. \\

@@

In the graphic interface there is the possibility to select the two set of five-year results to compare. The heatmap shows a colored square in a latitude and longitude quadrant if it's present a data for that quadrant in both the two sets of results. Till the period 1921-1925 there are few data in the results, with consequently few colored squares in the map. In the interface there is also the possibility to change the temperature metric that is used to compared the two set of results.\\
One problem of the interface is that the colored squares used in the map should change their height depending on the latitude value where they are positioned because the map projection isn't equiareal. The consequence in the heatmap visualization is that there is some empty space between the squares of a certain latitude value and the squares with the value of the latitude increased or decreased by one. This aspect makes the heatmap looks like as it is missing some data for certain latitude when instead the data are present, just the squares should be rectangles with a bigger height value.\\

By using the interactive map it is possible to notice an increase of avarage temperature in Europe and in Asia in the years from 1970 to the current years. In figure \ref{img:increaseMap} it is possible to see an example of increase between the two interval 1976/1980 to 2011/2015. The red colors are dominant in all the figure. Showing an increase of at least 2.5 degrees for all the regions that have a red square over them.

\begin{figure}[tbh]
\includegraphics[width=1\linewidth]{increaseEurope}
\caption{Map that shows the increase of the average temperature in Europe and in Asia between 1976 and 2015.}
\label{fig:interface}
\end{figure}

@@

\section{Conclusions}
\label{sec:con}
Before answering the question we asked earlier we will take a brief general overview of the characteristics of this project. One of the main problem we encountered is the limited (if not absence) of documentation describing the data. Especially for the older datasets we had to directly analyze the data to know the origin (region on earth) or the time coverage (day/night reports, coverage of all months). Moreover, we had to take into account during the analysis and the result discussion the anomalies we discovered in the station number charts (\ref{App:AppendixA}). Without a uniform and well formed source it's more difficult to extract reliable statistics.\\
We are overall satisfied with the performance of the program developed with Spark for analyzing the data. The program is able to analyze the entire dataset in approximately 8 hours. The program is also scalable to the data size. Future data in the next years will probably grow in dimensions, Spark can automatically handle new resources available in the cluster if there is the need. Furthermore the structure of the program allows also to manage the growth of the data by computing some intermediate results on the yearly data. The five year analysis is performed only on these intermediate results that are way smaller than the initial data. This feature removes the need to join all the data of the five years.\\
Other two aspect that improve the performance are the step we performed to group the big amount of files, and the caching of the data after the operations of parsing, filtering and grouping. The first improvement makes HDFS to be faster because it performs much better with a smaller amount of bigger files. The second improvement allows to compute different analysis on the same data without redoing some computation that are needed in the different types of analysis.\\

For what concerns the metrics used for analyzing the data, the average temperature is the most useful metric. It shows clearly the increase of the temperature in different periods and it's stable also if some particular events happen in short periods of the year. Minimum and maximum temperature are more susceptible to those events and indeed they show less clearly the increase of the temperature in the periods where the average metric reveals the increase. The standard deviation metric is useful to understand if there are errors in the data or if the data are not uniformly distributed (e.g. when the distributions of the stations is changed during the five year of the grouped results then the standard deviation results higher).
